---
title: "p8105_hw1_xy2517"
author: "Xuehan Yang Uni: xy2517"
date: "2021/9/22"
output: github_document
---


All the libraries needed in this project.

```{r message=FALSE}
library(tidyverse)
library(palmerpenguins)
```


# Problem 1

Create a dataframe comprised of numeric, logical, character, and factor variables.

```{r create a dataframe}
datatypes_df <-
  tibble(
    num_vec = rnorm(10,0,1), # numeric variable
    logic_vec = num_vec > 0, # logical variable
    char_vec = c("Stephen","Durant","James","Klay","CP3","Davis","Giannis","Joker","Harden","Irving"), # character variable
    factor_vec = factor(c("S","S","A","A","B","A","S","A","S","A"), levels = c("B","A","S")) # factor variables and order them
  )
```

Take the Mean of each variable of the dataframe.

```{r calculate the means, results='hide', warning=FALSE}
library(tidyverse)
mean(pull(datatypes_df, 1)) # try the mean for numeric variable
mean(pull(datatypes_df, 2)) # try the mean for logical variable
mean(pull(datatypes_df, 3)) # try the mean for character variable
mean(pull(datatypes_df, 4)) # try the mean for factor variable
```

From the results, I can see that only numeric and logical variables can be put in the mean() function. 

Then convert logical, character and factor variables to numeric.

```{r convert to numeric variable, results='hide',warning=FALSE}
as.numeric(pull(datatypes_df,2))
as.numeric(pull(datatypes_df,3))
as.numeric(pull(datatypes_df,4))
```

Conclusions:

Logical and factor variales can be converted to numeric with function as.numeric, however, character variables can't.

* For logical variables, "TRUE" equals to 1 and "FALSE" equals to 0.

* For character variables, they cannot be converted to numeric.

* For factor variables, they can be converted to numeric according to the alphabetical order by default 
unless you give them order at the beginning.


# Problem 2

After installing packages "palmerpenguins", load dataset "penguins".

```{r load penguins dataset}
data("penguins", package = "palmerpenguins")
```

Then, try to get some descriptive statistics of this dataset.

## Description 1: Names/values of important variables


```{r values of important variables}
head(penguins,5)
penguins = drop_na(penguins)
```

Drop the NA values.

As we can see, this dataset contains some main characteristics of penguins which includes `r names(penguins)`. There are `r ncol(penguins)` columns and 3 factors, 3int and 3 doubles types of data are included.

There are three penguin species and three living islands. 
The average bill_length is `r round(mean(pull(penguins,bill_length_mm)),0)`mm, 
average bill_depth is `r round(mean(pull(penguins,bill_depth_mm)),0)`mm, 
average flipper_length is `r round(mean(pull(penguins,flipper_length_mm)),0)`mm,
average body_mass is `r round(mean(pull(penguins,body_mass_g)),0)`g. 
`r sum(penguins$sex=="female")` of them are female and `r sum(penguins$sex=="male")` of them are male. 

## Description 2: Size of the dataset

There are `r nrow(penguins)` rows and `r ncol(penguins)` columns in this dataset.

## Description 3: Relation between bill_length and flipper_length

```{r warning=FALSE}
ggplot(data = penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point() + 
  ggtitle("Relation between bill_length and flipper_length") + 
  theme(plot.title = element_text(hjust = 0.5))

ggsave("./plots/bill_length_flipper_length.pdf", height = 7, width = 8)
```


