---
title: "p8105_hw1_xy2517"
author: "Xuehan Yang Uni: xy2517"
date: "2021/9/22"
output: github_document
---


All the libraries needed in this project.

```{r message=FALSE}
library(tidyverse)
library(palmerpenguins)
```


# Problem 1

Create a dataframe comprised of numeric, logical, character, and factor variables.

```{r create a dataframe}
datatypes_df <-
  tibble(
    num_vec = rnorm(10,0,1), # numeric variable
    logic_vec = num_vec > 0, # logical variable
    char_vec = c("Stephen","Durant","James","Klay","CP3","Davis","Giannis","Joker","Harden","Irving"), # character variable
    factor_vec = factor(c("S","S","A","A","B","A","S","A","S","A"), levels = c("B","A","S")) # factor variables and order them
  )
```

Take the Mean of each variable of the dataframe.

```{r calculate the means, results='hide', warning=FALSE}
library(tidyverse)
mean(pull(datatypes_df, 1)) # try the mean for numeric variable
mean(pull(datatypes_df, 2)) # try the mean for logical variable
mean(pull(datatypes_df, 3)) # try the mean for character variable
mean(pull(datatypes_df, 4)) # try the mean for factor variable
```

From the results, I can see that only numeric and logical variables can be put in the mean() function. 

Then convert logical, character and factor variables to numeric.

```{r convert to numeric variable, results='hide',warning=FALSE}
as.numeric(pull(datatypes_df,2))
as.numeric(pull(datatypes_df,3))
as.numeric(pull(datatypes_df,4))
```

Conclusions:

Logical and factor variales can be converted to numeric with function as.numeric, however, character variables can't.

* For logical variables, "TRUE" equals to 1 and "FALSE" equals to 0.

* For character variables, they cannot be converted to numeric.

* For factor variables, they can be converted to numeric according to the alphabetical order by default 
unless you give them order at the beginning.


# Problem 2

After installing packages "palmerpenguins", load dataset "penguins".

```{r load penguins dataset}
data("penguins", package = "palmerpenguins")
```

Then, try to get some descriptive statistics of this dataset.

## Description 1: Names/values of important variables

```{r get names}
names(penguins) 
```

```{r values of important variables}
head(penguins,5)
```

As we can see, this dataset contains some main characteristics of penguins like species, body_mass_g and body shape indexes. There are 8 columns and 3 factors, 3int and 3 doubles types of data are included.

```{r view discriptive statistics}
summary(penguins)
```

There are three penguin species and three living islands. The average bill_length is 43.92mm, average bill_depth is 17.15mm, average flipper_length is 200.9mm and average body_mass is 4202g. 165 of them are female and 168 of them are male. 

There are also some NA values in dataset. If we are going to do statistical inference, data manipulation should be done here.

## Description 2: Size of the dataset

```{r get row and col numbers}
nrow(penguins)
ncol(penguins)
```

There are 344 rows and 8 columns in this dataset.

## Description 3: Relation between bill_length and flipper_length

```{r warning=FALSE}
ggplot(data = penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point() + labs(title = "Relation between bill_length and flipper_length")
```


